rbcL ASV clustering

### 2019-11-02 code/rbcl_asv_clustering.Rmd
Chloromonas krienitzii not really an OTU, change labelling system

Manually edit the clustering:
  - *Chloromonas krienitzii* gets its own cluster, since this is highly abundant

```{r edit-otus, echo=F}
tsne_edited_otus <- tsne_taxa_otu %>%
  mutate(otu = case_when((V2 < 14  & V2 > 8 & V1 > (-20) & V1 < (-12)) ~ "9",# krienitzii
                         TRUE ~ otu))

@@

# name each OTU, lettered in order of rel abund
otu_names <- tibble(otu = as.character(1:9), 
                    otu_name = c("Chloromonas spp. C",
                                 "Sanguina B",
                                 "Chlainomonas A",
                                 "Chloromonas muramotoi E",
                                 "Chloromonas F",
                                 "Chloromonas H",
                                 "Algae G",
                                 "Trebouxiophyceae I",
                                 "Chloromonas krienitzii D"))

@@

final_tsne <- ggplot(tsne_rel_abund, aes(x=V1, y=V2)) +
  ggforce::geom_mark_ellipse(aes(fill=otu_name, label=otu_name), 
                             show.legend = FALSE, 
                             alpha=0.2, 
                             linetype="blank", 
                             expand = unit(2, "mm")) +
  # plot points not assigned to genus or species
  geom_point(data = filter(tsne_rel_abund, is.na(plot_taxa)), 
             aes(size = global_rel_abund),
             color = "grey",alpha = 0.5) +
  # color points that were assigned to genus or species
  geom_point(data = filter(tsne_rel_abund, !is.na(plot_taxa)), 
             aes(color = plot_taxa %>% fct_relevel(taxa.order),
                 size = global_rel_abund), 
             alpha = 0.5) +
  labs(color="ASV taxonomic assignment") +
  ggthemes::theme_few() +
  scale_size(range = c(1, 6), guide="none") + ################################################# not changing the scale size
  theme(legend.position = "none")
final_tsne
```

@@

mds_asv_plot <- mds_coords %>% 
  left_join(taxonomy, by="asv_id") %>% 
  left_join(asv_otu_key, by="asv_id") %>% 
  left_join(total_rel_abund, by="asv_id") %>% 
  filter(!(asv_id %in% c("639", "534", "646"))) # remove two trebouxiophyceae outliers

mds <- ggplot(mds_asv_plot, aes(V1, V2)) + 
  ggforce::geom_mark_ellipse(data = mds_asv_plot %>% 
                               filter(otu_name != "Trebouxiophyceae J"), # dont plot the ellipse around other treb
                             aes(fill = otu_name, label=otu_name),
                             show.legend = FALSE, 
                             alpha=0.2, 
                             linetype="blank",
                             expand = unit(2, "mm")) +
  geom_point(aes(alpha = 0.7, color = plot_taxa, size = global_rel_abund)) +
  scale_size(range = c(1, 6), guide="none") +
  scale_alpha(guide="none")+
  theme_few() +
  labs(color="ASV taxonomic assignment")
  # geom_text(aes(label = asv_id)) # too see outliers, asv_id=c("639", "534")
mds
```

@@

# TO do
- add MDS
- point size proportional to relative abundance
- ggtree
- show different values of perplexity yield comparable results

>>>>

Manually edit the clustering to show *Chloromonas krienitzii* on relative abundance barplot

```{r edit-otus, echo=F}
tsne_edited_otus <- tsne_taxa_otu %>%
  mutate(otu = case_when((plot_taxa == "Chloromonas_krienitzii") ~ "9",
                         TRUE ~ otu))

@@

# name each OTU, lettered in order of rel abund
otu_names <- tibble(otu = as.character(1:9), 
                    otu_name = c("Chloromonas spp. C",
                                 "Sanguina B",
                                 "Chlainomonas A",
                                 "Chloromonas muramotoi D",
                                 "Chloromonas E",
                                 "Chloromonas G",
                                 "Algae F",
                                 "Trebouxiophyceae H",
                                 "Chloromonas krienitzii"))

@@

final_tsne <- ggplot(tsne_rel_abund, aes(x=V1, y=V2)) +
  ggforce::geom_mark_ellipse(aes(fill=otu_name, label=otu_name), 
                             show.legend = FALSE, 
                             alpha=0.2, 
                             linetype="blank", 
                             expand = unit(2, "mm")) +
  # plot points not assigned to genus or species
  geom_point(data = filter(tsne_rel_abund, is.na(plot_taxa)), 
             aes(size = global_rel_abund),
             color = "grey",alpha = 0.5) +
  # color points that were assigned to genus or species
  geom_point(data = filter(tsne_rel_abund, !is.na(plot_taxa)), 
             aes(color = plot_taxa %>% fct_relevel(taxa.order),
                 size = global_rel_abund), 
             alpha = 0.5) +
  labs(color="ASV taxonomic assignment", tag="A") +
  ggthemes::theme_few() +
  scale_size(range = c(1, 6), guide="none") + ################################################# not changing the scale size
  theme(legend.position = "none")
final_tsne
```

@@

mds_asv_plot <- mds_coords %>% 
  left_join(taxonomy, by="asv_id") %>% 
  left_join(asv_otu_key, by="asv_id") %>% 
  left_join(total_rel_abund, by="asv_id") %>% 
  filter(!(asv_id %in% c("639", "534", "646"))) # remove Trebouxiophyceae outliers

mds <- ggplot(mds_asv_plot, aes(V1, V2)) + 
  ggforce::geom_mark_ellipse(data = mds_asv_plot %>% 
                               filter(otu_name != "Trebouxiophyceae J"), # dont plot the ellipse around other treb
                             aes(fill = otu_name, label=otu_name),
                             show.legend = FALSE, 
                             alpha=0.2, 
                             linetype="blank",
                             expand = unit(2, "mm")) +
  geom_point(aes(alpha = 0.7, color = plot_taxa, size = global_rel_abund)) +
  scale_size(range = c(1, 6), guide="none") +
  scale_alpha(guide="none")+
  theme_few() +
  labs(color="ASV taxonomic assignment", tag="B")
  # geom_text(aes(label = asv_id)) # too see outliers, asv_id=c("639", "534")
mds
```

@@
